---
title: "Hurricane Ian (2022)"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-ian
library(tidyverse)
library(patchwork)

# Load hourly KTPA observations
event_data <- read_csv("data/ktpa_2022.csv", col_types = cols(.default = col_character()), show_col_types = FALSE) |>
  filter(grepl(":53:", DATE)) |>
  mutate(
    datetime = as.POSIXct(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  ) |>
  filter(datetime >= as.POSIXct("2022-09-26 00:00:00") & datetime <= as.POSIXct("2022-09-29 23:59:59"))
```

::: {.news-header}
<span class="news-badge">CATEGORY 4/5 NEAR-MISS & NEGATIVE SURGE</span>
<h1 style="color: #ffffff; margin-top: 0.5rem; margin-bottom: 0.5rem; font-size: 2.2rem; font-weight: 800;">Hurricane Ian (September 2022)</h1>
<div class="news-station-tag">STATION: KTPA &bull; MASSIVE DRAIN OF TAMPA BAY BED</div>
:::

::: {.records-grid}
::: {.record-card}
<div class="record-icon">📉</div>
<div class="record-value">29.56 inHg</div>
<div class="record-label">Lowest Pressure</div>
<div class="record-sub">1001.0 hPa on Sep 28</div>
:::

::: {.record-card}
<div class="record-icon">💨</div>
<div class="record-value">30.0 mph</div>
<div class="record-label">Peak Sustained Wind</div>
<div class="record-sub">Gusts to 50+ mph in squalls</div>
:::

::: {.record-card}
<div class="record-icon">🌧️</div>
<div class="record-value">1.83 in</div>
<div class="record-label">Storm Rainfall</div>
<div class="record-sub">Track shifted south of bay</div>
:::

::: {.record-card}
<div class="record-icon">🌊</div>
<div class="record-value">-5.5 ft Blowout</div>
<div class="record-label">Seabed Exposure</div>
<div class="record-sub">Historic negative surge</div>
:::
:::

```{r}
#| label: fig-telemetry-ian
#| fig-cap: "Hourly meteorological telemetry recorded at Tampa International Airport (KTPA) during the event window. Top: Barometric pressure (inHg). Bottom: Sustained wind speed (mph)."
#| fig-width: 9.5
#| fig-height: 5.6
#| dpi: 180

# Panel 1: Barometric Pressure
p1 <- ggplot(event_data, aes(x = datetime, y = pressure)) +
  geom_line(color = "#7e22ce", linewidth = 1.1) +
  geom_point(data = filter(event_data, pressure == min(pressure, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    title = "Hurricane Ian (2022): KTPA Telemetry Breakdown",
    subtitle = "Station Pressure (inHg) & Sustained Wind Speed (mph)",
    y = "Pressure (inHg)",
    x = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 13, color = "#0f172a"),
    plot.subtitle = element_text(color = "#475569", size = 10, margin = margin(b = 8)),
    axis.title.y = element_text(face = "bold", color = "#7e22ce", size = 10),
    panel.grid.minor = element_blank()
  )

# Panel 2: Wind Speed
p2 <- ggplot(event_data, aes(x = datetime, y = wind)) +
  geom_line(color = "#6b21a8", linewidth = 1.1) +
  geom_point(data = filter(event_data, wind == max(wind, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    y = "Wind Speed (mph)",
    x = "Date & Time (EDT)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.title.x = element_text(face = "bold", color = "#0f172a", margin = margin(t = 6)),
    axis.title.y = element_text(face = "bold", color = "#6b21a8", size = 10),
    panel.grid.minor = element_blank()
  )

p1 / p2
```


## Synoptic Evolution & Last-Minute Track Shift

In late September 2022, Hurricane Ian intensified into a monstrous high-end Category 4/5 cyclone over the southeastern Gulf. Initial NHC forecast tracks pointed toward a catastrophic worst-case direct landfall inside Tampa Bay, triggering mandatory evacuations of over 300,000 residents across Hillsborough and Pinellas counties.

On the morning of September 28, Ian executed an eastward turn, making landfall 85 miles south at Cayo Costa / Fort Myers with 150+ mph winds. While Tampa Bay was spared a catastrophic direct surge, the southern landfall track created a historic meteorological phenomenon.

## Meteorological Telemetry at KTPA

Because Ian made landfall to the south, Tampa Bay sat squarely on the cold, offshore northern quadrant of the storm:

- **Extreme Negative Surge:** Northerly and northeasterly winds blowing between **30 and 45 mph** literally siphoned water out of Tampa Bay, lowering water levels to **5.5 feet below normal**. Hillsborough Bay, Bayshore Boulevard, and Davis Island basins became barren mudflats where residents famously walked on the dry bay floor.
- **Pressure Drop:** Station pressure dropped to **29.56 inHg** at KTPA on the evening of September 28 as Ian crossed the southern peninsula.
- **Spared Catastrophic Deluge:** Because Ian's core rain shield drifted across central and southern Florida, KTPA received **1.83 inches of rain**, escaping the 15–20 inch deluges recorded just miles to the south in Sarasota and Lee counties.


---

::: {.callout-note}
### Explore More Weather Crises
Return to the [Natural Disasters Overview](disasters.qmd) to view the full 10-year timeline, comparison tables, and summaries of all major storms.
:::
