---
title: "Hurricane Michael (2018)"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-michael
library(tidyverse)
library(patchwork)

# Load hourly KTPA observations
event_data <- read_csv("data/ktpa_2018.csv", col_types = cols(.default = col_character()), show_col_types = FALSE) |>
  filter(grepl(":53:", DATE)) |>
  mutate(
    datetime = as.POSIXct(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  ) |>
  filter(datetime >= as.POSIXct("2018-10-08 00:00:00") & datetime <= as.POSIXct("2018-10-11 23:59:59"))
```

::: {.news-header}
<span class="news-badge">CATEGORY 5 SYSTEM PASSAGE</span>
<h1 style="color: #ffffff; margin-top: 0.5rem; margin-bottom: 0.5rem; font-size: 2.2rem; font-weight: 800;">Hurricane Michael (October 2018)</h1>
<div class="news-station-tag">STATION: KTPA &bull; GULF TRANSIT & REVERSE NEGATIVE SURGE</div>
:::

::: {.records-grid}
::: {.record-card}
<div class="record-icon">📉</div>
<div class="record-value">29.75 inHg</div>
<div class="record-label">Lowest Pressure</div>
<div class="record-sub">1007.4 hPa on Oct 10</div>
:::

::: {.record-card}
<div class="record-icon">💨</div>
<div class="record-value">22.0 mph</div>
<div class="record-label">Peak Sustained Wind</div>
<div class="record-sub">Gusts to 33 mph offshore flow</div>
:::

::: {.record-card}
<div class="record-icon">🌧️</div>
<div class="record-value">0.50 in</div>
<div class="record-label">Storm Rainfall</div>
<div class="record-sub">Dry offshore slotting</div>
:::

::: {.record-card}
<div class="record-icon">🌊</div>
<div class="record-value">-2.5 ft Drain</div>
<div class="record-label">Bay Dynamics</div>
<div class="record-sub">Water siphoned into Gulf</div>
:::
:::

```{r}
#| label: fig-telemetry-michael
#| fig-cap: "Hourly meteorological telemetry recorded at Tampa International Airport (KTPA) during the event window. Top: Barometric pressure (inHg). Bottom: Sustained wind speed (mph)."
#| fig-width: 9.5
#| fig-height: 5.6
#| dpi: 180

# Panel 1: Barometric Pressure
p1 <- ggplot(event_data, aes(x = datetime, y = pressure)) +
  geom_line(color = "#0369a1", linewidth = 1.1) +
  geom_point(data = filter(event_data, pressure == min(pressure, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    title = "Hurricane Michael (2018): KTPA Telemetry Breakdown",
    subtitle = "Station Pressure (inHg) & Sustained Wind Speed (mph)",
    y = "Pressure (inHg)",
    x = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 13, color = "#0f172a"),
    plot.subtitle = element_text(color = "#475569", size = 10, margin = margin(b = 8)),
    axis.title.y = element_text(face = "bold", color = "#0369a1", size = 10),
    panel.grid.minor = element_blank()
  )

# Panel 2: Wind Speed
p2 <- ggplot(event_data, aes(x = datetime, y = wind)) +
  geom_line(color = "#0284c7", linewidth = 1.1) +
  geom_point(data = filter(event_data, wind == max(wind, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    y = "Wind Speed (mph)",
    x = "Date & Time (EDT)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.title.x = element_text(face = "bold", color = "#0f172a", margin = margin(t = 6)),
    axis.title.y = element_text(face = "bold", color = "#0284c7", size = 10),
    panel.grid.minor = element_blank()
  )

p1 / p2
```


## Synoptic Evolution & Gulf Transit

In early October 2018, Hurricane Michael underwent explosive rapid intensification over the warm waters of the eastern Gulf of Mexico, deepening into an historic Category 5 hurricane with 160 mph sustained winds before devastating Mexico Beach and the Florida Panhandle.

While Michael's center remained 200 miles west of Tampa as it rocketed north toward the Panhandle, its enormous wind field and intense low pressure exerted a profound hydrodynamic impact on Tampa Bay's coastal waters.

## Meteorological Telemetry at KTPA

Despite Michael's core remaining far offshore, KTPA's automated sensors clearly chronicled the cyclone's passage:

- **Broad Pressure Deficit:** Station pressure dipped to **29.75 inHg** on October 10 as Michael reached peak Category 5 intensity off the Florida coast.
- **Reverse Negative Storm Surge:** As Michael churned to the west, its counterclockwise winds drove steady northeasterly and easterly breezes across Tampa Bay. This continuous offshore gale pushed bay water toward the Gulf, lowering water levels across Hillsborough Bay by **2 to 2.5 feet below normal astronomical tide**.
- **Dry Slotting:** Unlike Hermine or Debby, Tampa Bay found itself in a continental dry slot behind Michael's eastern rainbands, holding storm rainfall to just **0.50 inches** at KTPA.


---

::: {.callout-note}
### Explore More Weather Crises
Return to the [Natural Disasters Overview](disasters.qmd) to view the full 10-year timeline, comparison tables, and summaries of all major storms.
:::
