---
title: "Hurricane Idalia (2023)"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-idalia
library(tidyverse)
library(patchwork)

# Load hourly KTPA observations
event_data <- read_csv("data/ktpa_2023.csv", col_types = cols(.default = col_character()), show_col_types = FALSE) |>
  filter(grepl(":53:", DATE)) |>
  mutate(
    datetime = as.POSIXct(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  ) |>
  filter(datetime >= as.POSIXct("2023-08-28 00:00:00") & datetime <= as.POSIXct("2023-08-31 23:59:59"))
```

::: {.news-header}
<span class="news-badge">HIGH-TIDE SURGE INUNDATION</span>
<h1 style="color: #ffffff; margin-top: 0.5rem; margin-bottom: 0.5rem; font-size: 2.2rem; font-weight: 800;">Hurricane Idalia (August 2023)</h1>
<div class="news-station-tag">STATION: KTPA &bull; KING TIDE SURGE & RECORD BAYSHORE INUNDATION</div>
:::

::: {.records-grid}
::: {.record-card}
<div class="record-icon">📉</div>
<div class="record-value">29.55 inHg</div>
<div class="record-label">Lowest Pressure</div>
<div class="record-sub">1000.7 hPa on Aug 30</div>
:::

::: {.record-card}
<div class="record-icon">💨</div>
<div class="record-value">33.0 mph</div>
<div class="record-label">Peak Sustained Wind</div>
<div class="record-sub">Gusts to 52 mph onshore</div>
:::

::: {.record-card}
<div class="record-icon">🌧️</div>
<div class="record-value">2.68 in</div>
<div class="record-label">Storm Rainfall</div>
<div class="record-sub">Outer squall bands</div>
:::

::: {.record-card}
<div class="record-icon">🌊</div>
<div class="record-value">4.0 – 5.0 ft</div>
<div class="record-label">Peak Storm Surge</div>
<div class="record-sub">Inundated Bayshore & Davis Is.</div>
:::
:::

```{r}
#| label: fig-telemetry-idalia
#| fig-cap: "Hourly meteorological telemetry recorded at Tampa International Airport (KTPA) during the event window. Top: Barometric pressure (inHg). Bottom: Sustained wind speed (mph)."
#| fig-width: 9.5
#| fig-height: 5.6
#| dpi: 180

# Panel 1: Barometric Pressure
p1 <- ggplot(event_data, aes(x = datetime, y = pressure)) +
  geom_line(color = "#047857", linewidth = 1.1) +
  geom_point(data = filter(event_data, pressure == min(pressure, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    title = "Hurricane Idalia (2023): KTPA Telemetry Breakdown",
    subtitle = "Station Pressure (inHg) & Sustained Wind Speed (mph)",
    y = "Pressure (inHg)",
    x = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 13, color = "#0f172a"),
    plot.subtitle = element_text(color = "#475569", size = 10, margin = margin(b = 8)),
    axis.title.y = element_text(face = "bold", color = "#047857", size = 10),
    panel.grid.minor = element_blank()
  )

# Panel 2: Wind Speed
p2 <- ggplot(event_data, aes(x = datetime, y = wind)) +
  geom_line(color = "#059669", linewidth = 1.1) +
  geom_point(data = filter(event_data, wind == max(wind, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    y = "Wind Speed (mph)",
    x = "Date & Time (EDT)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.title.x = element_text(face = "bold", color = "#0f172a", margin = margin(t = 6)),
    axis.title.y = element_text(face = "bold", color = "#059669", size = 10),
    panel.grid.minor = element_blank()
  )

p1 / p2
```


## Synoptic Evolution & Gulf Acceleration

In late August 2023, Hurricane Idalia rapidly intensified over exceptionally warm eastern Gulf waters, ballooning into a Category 4 major hurricane before making landfall in Florida's Big Bend as a Category 3 storm on August 30.

Although Idalia tracked roughly 120 miles west of Tampa Bay, its eastern flank unleashed persistent southerly and southwesterly gales that pushed Gulf water directly into the mouth of Tampa Bay. Crucially, the surge arrived precisely during an astronomical **King Tide** (supermoon peak).

## Meteorological Telemetry at KTPA

KTPA's telemetry documented the dynamics of a high-tide surge disaster:

- **Pressure & Wind Signature:** Station pressure fell to **29.55 inHg** on August 30, with sustained onshore winds of **33 mph** and squall gusts over 50 mph driving ocean water inland.
- **Unprecedented Coastal Inundation:** The combined astronomical King Tide and 4 to 5 foot surge pushed Tampa Bay to its highest water level since 1993 at the time. Water completely submerged Bayshore Boulevard, flowed through ground-floor condominiums on Davis Islands, and cut off access to Gulf beach communities in St. Petersburg and Clearwater.
- **Rainfall Total:** Outer bands dropped **2.68 inches of rain** at KTPA, while the primary damage was caused by saltwater flooding.


---

::: {.callout-note}
### Explore More Weather Crises
Return to the [Natural Disasters Overview](disasters.qmd) to view the full 10-year timeline, comparison tables, and summaries of all major storms.
:::
