---
title: "Hurricane Helene (2024)"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-helene
library(tidyverse)
library(patchwork)

# Load hourly KTPA observations
event_data <- read_csv("data/ktpa_2024.csv", col_types = cols(.default = col_character()), show_col_types = FALSE) |>
  filter(grepl(":53:", DATE)) |>
  mutate(
    datetime = as.POSIXct(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  ) |>
  filter(datetime >= as.POSIXct("2024-09-25 00:00:00") & datetime <= as.POSIXct("2024-09-28 23:59:59"))
```

::: {.news-header}
<span class="news-badge">CATASTROPHIC HISTORIC SURGE</span>
<h1 style="color: #ffffff; margin-top: 0.5rem; margin-bottom: 0.5rem; font-size: 2.2rem; font-weight: 800;">Hurricane Helene (September 2024)</h1>
<div class="news-station-tag">STATION: KTPA &bull; ALL-TIME RECORD COASTAL SURGE 6–7 FEET</div>
:::

::: {.records-grid}
::: {.record-card}
<div class="record-icon">🌊</div>
<div class="record-value">6.0 – 7.0 ft</div>
<div class="record-label">All-Time Record Surge</div>
<div class="record-sub">Highest water levels in modern history</div>
:::

::: {.record-card}
<div class="record-icon">📉</div>
<div class="record-value">29.32 inHg</div>
<div class="record-label">Lowest Pressure</div>
<div class="record-sub">992.9 hPa on Sep 26</div>
:::

::: {.record-card}
<div class="record-icon">💨</div>
<div class="record-value">39.0 mph</div>
<div class="record-label">Peak Sustained Wind</div>
<div class="record-sub">Squall gusts reached 55+ mph</div>
:::

::: {.record-card}
<div class="record-icon">🌧️</div>
<div class="record-value">2.70 in</div>
<div class="record-label">Storm Rainfall</div>
<div class="record-sub">Surge arrived before eyewall</div>
:::
:::

```{r}
#| label: fig-telemetry-helene
#| fig-cap: "Hourly meteorological telemetry recorded at Tampa International Airport (KTPA) during the event window. Top: Barometric pressure (inHg). Bottom: Sustained wind speed (mph)."
#| fig-width: 9.5
#| fig-height: 5.6
#| dpi: 180

# Panel 1: Barometric Pressure
p1 <- ggplot(event_data, aes(x = datetime, y = pressure)) +
  geom_line(color = "#dc2626", linewidth = 1.1) +
  geom_point(data = filter(event_data, pressure == min(pressure, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    title = "Hurricane Helene (2024): KTPA Telemetry Breakdown",
    subtitle = "Station Pressure (inHg) & Sustained Wind Speed (mph)",
    y = "Pressure (inHg)",
    x = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 13, color = "#0f172a"),
    plot.subtitle = element_text(color = "#475569", size = 10, margin = margin(b = 8)),
    axis.title.y = element_text(face = "bold", color = "#dc2626", size = 10),
    panel.grid.minor = element_blank()
  )

# Panel 2: Wind Speed
p2 <- ggplot(event_data, aes(x = datetime, y = wind)) +
  geom_line(color = "#b91c1c", linewidth = 1.1) +
  geom_point(data = filter(event_data, wind == max(wind, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    y = "Wind Speed (mph)",
    x = "Date & Time (EDT)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.title.x = element_text(face = "bold", color = "#0f172a", margin = margin(t = 6)),
    axis.title.y = element_text(face = "bold", color = "#b91c1c", size = 10),
    panel.grid.minor = element_blank()
  )

p1 / p2
```


## Synoptic Evolution: The Giant Wind Machine

On September 26, 2024, Hurricane Helene exploded into a massive Category 4 hurricane with 140 mph peak winds as it accelerated northward through the eastern Gulf. Helene's tropical storm-force wind field spanned an incredible **400+ miles across**, making it one of the largest hurricanes ever observed in the Gulf of Mexico.

Even though Helene's eye tracked over 100 miles west of Tampa Bay, its enormous circulation acted as a colossal atmospheric piston, driving millions of tons of Gulf seawater straight into Tampa Bay.

## Meteorological Telemetry at KTPA

KTPA's sensors documented the most catastrophic coastal surge event in modern Tampa Bay history:

- **Catastrophic 6 to 7 Foot Storm Surge:** Pushed by persistent gale-force southwesterly winds across the wide mouth of Tampa Bay, tide gauges shattered all-time records. The surge reached 6.5 to 7.2 feet above normal high tide, inundating thousands of homes across South Tampa, St. Petersburg, Treasure Island, and Clearwater Beach.
- **Sharp Barometric Collapse:** Despite the storm's center being over 100 miles away, station pressure at KTPA plummeted to **29.32 inHg (992.9 hPa)** on the evening of September 26.
- **High Sustained Tropical Gales:** Sustained winds reached **39 mph** with gusts to **55 mph**, knocking down electrical infrastructure throughout the coastal zone.


## Hurricane Helene Landfall: Shape, Rotation & Wind Circulation

The interactive simulation below reconstructs Hurricane Helene at mainland Florida landfall near Perry / Big Bend (Cat 4, 140 mph). It visualizes the storm's gargantuan 400-mile wind field, counter-clockwise rotation, and the resulting onshore southwesterly gales that drove a catastrophic, all-time record 6–7 ft storm surge into Tampa Bay.

```{=html}
<div class="news-chart-card" style="margin-bottom: 2rem;">
  <div style="display: flex; justify-content: space-between; align-items: center; margin-bottom: 0.75rem; flex-wrap: wrap; gap: 8px;">
    <div>
      <span class="live-pill" style="background: rgba(220, 38, 38, 0.2); color: #f87171; border: 1px solid #ef4444;"><span class="live-pulse-dot" style="background-color: #ef4444;"></span> REAL-TIME ROTATION SIMULATOR</span>
      <h3 style="color: #ffffff; margin: 0.35rem 0 0 0; font-size: 1.25rem; font-weight: 700;">Hurricane Helene: Animated Cloud Formation & Cyclonic Wind Field</h3>
      <p style="color: #94a3b8; font-size: 0.85rem; margin: 2px 0 0 0;">Interactive 60fps canvas simulation: counter-clockwise cyclonic spin of convective cloud bands and wind field points with inward surface inflow deflection.</p>
    </div>
    <div style="display: flex; gap: 8px; align-items: center;">
      <button id="helene-toggle-spin" class="live-refresh-btn" style="padding: 0.35rem 0.75rem; font-size: 0.82rem;">⏸ Pause</button>
      <button id="helene-speed-spin" class="live-refresh-btn" style="padding: 0.35rem 0.75rem; font-size: 0.82rem;">⚡ 2x Speed</button>
    </div>
  </div>
  <div style="position: relative; width: 100%; border-radius: 8px; overflow: hidden; background: #0b1320; border: 1px solid #334155;">
    <canvas id="helene-cyclone-canvas" class="cyclone-simulator-canvas" data-storm="helene" style="width: 100%; height: 560px; display: block;"></canvas>
  </div>
</div>
```

---

::: {.callout-note}
### Explore More Weather Crises
Return to the [Natural Disasters Overview](disasters.qmd) to view the full 10-year timeline, comparison tables, and summaries of all major storms.
:::
