---
title: "Hurricane Milton (2024)"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-milton
library(tidyverse)
library(patchwork)

# Load hourly KTPA observations
event_data <- read_csv("data/ktpa_2024.csv", col_types = cols(.default = col_character()), show_col_types = FALSE) |>
  filter(grepl(":53:", DATE)) |>
  mutate(
    datetime = as.POSIXct(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  ) |>
  filter(datetime >= as.POSIXct("2024-10-08 00:00:00") & datetime <= as.POSIXct("2024-10-11 23:59:59"))
```

::: {.news-header}
<span class="news-badge">DIRECT STRIKE CAT 3 LANDFALL</span>
<h1 style="color: #ffffff; margin-top: 0.5rem; margin-bottom: 0.5rem; font-size: 2.2rem; font-weight: 800;">Hurricane Milton (October 2024)</h1>
<div class="news-station-tag">STATION: KTPA &bull; DIRECT EYEWALL PASSAGE & 16.03 INCHES RAIN</div>
:::

::: {.records-grid}
::: {.record-card}
<div class="record-icon">🌀</div>
<div class="record-value">Direct Strike</div>
<div class="record-label">Landfall Proximity</div>
<div class="record-sub">Landfall Siesta Key, eyewall over bay</div>
:::

::: {.record-card}
<div class="record-icon">📉</div>
<div class="record-value">29.18 inHg</div>
<div class="record-label">Lowest Pressure</div>
<div class="record-sub">988.2 hPa inside eyewall</div>
:::

::: {.record-card}
<div class="record-icon">🌧️</div>
<div class="record-value">16.03 in</div>
<div class="record-label">Eyewall Deluge</div>
<div class="record-sub">Catastrophic 24-hr deluge</div>
:::

::: {.record-card}
<div class="record-icon">💨</div>
<div class="record-value">60 – 100 mph</div>
<div class="record-label">Hurricane Gusts</div>
<div class="record-sub">Tore roof off Tropicana Field</div>
:::
:::

```{r}
#| label: fig-telemetry-milton
#| fig-cap: "Hourly meteorological telemetry recorded at Tampa International Airport (KTPA) during the event window. Top: Barometric pressure (inHg). Bottom: Sustained wind speed (mph)."
#| fig-width: 9.5
#| fig-height: 5.6
#| dpi: 180

# Panel 1: Barometric Pressure
p1 <- ggplot(event_data, aes(x = datetime, y = pressure)) +
  geom_line(color = "#991b1b", linewidth = 1.1) +
  geom_point(data = filter(event_data, pressure == min(pressure, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    title = "Hurricane Milton (2024): KTPA Telemetry Breakdown",
    subtitle = "Station Pressure (inHg) & Sustained Wind Speed (mph)",
    y = "Pressure (inHg)",
    x = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 13, color = "#0f172a"),
    plot.subtitle = element_text(color = "#475569", size = 10, margin = margin(b = 8)),
    axis.title.y = element_text(face = "bold", color = "#991b1b", size = 10),
    panel.grid.minor = element_blank()
  )

# Panel 2: Wind Speed
p2 <- ggplot(event_data, aes(x = datetime, y = wind)) +
  geom_line(color = "#7f1d1d", linewidth = 1.1) +
  geom_point(data = filter(event_data, wind == max(wind, na.rm = TRUE)), color = "#dc2626", size = 3) +
  labs(
    y = "Wind Speed (mph)",
    x = "Date & Time (EDT)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.title.x = element_text(face = "bold", color = "#0f172a", margin = margin(t = 6)),
    axis.title.y = element_text(face = "bold", color = "#7f1d1d", size = 10),
    panel.grid.minor = element_blank()
  )

p1 / p2
```


## Synoptic Evolution: The Direct Hit

Just twelve days after Hurricane Helene devastated Tampa's coastlines, Hurricane Milton formed in the southwestern Gulf and underwent one of the fastest explosive intensifications in world history, dropping to **897 hPa with 180 mph winds** as a Category 5 superstorm.

Milton made landfall on the evening of October 9, 2024 near Siesta Key / Sarasota as a Category 3 hurricane. Because Milton made landfall immediately south of Tampa Bay, its northern eyewall—the most dangerous portion of the storm—slammed directly across St. Petersburg, Clearwater, Tampa, and KTPA.

## Meteorological Telemetry at KTPA

Milton subjected Tampa Bay to simultaneous extreme rainfall and destructive eyewall winds:

- **Historic 16.03-Inch Deluge:** The northern eyewall stalled over northern Pinellas and Hillsborough counties, dumping **16.03 inches of rain** at KTPA in under 24 hours. This caused catastrophic flash flooding that submerged neighborhoods miles inland.
- **Eyewall Barometric Depression:** Station pressure plummeted to **29.18 inHg (988.2 hPa)** as the eyewall passed directly over the station.
- **Destructive Wind Force:** Extreme hurricane-force gusts shredded trees, flattened power grids across the entire region, and famously shredded the fiberglass roof of Tropicana Field in downtown St. Petersburg.


## Hurricane Milton Landfall: Shape, Rotation & Wind Circulation

The interactive simulation below reconstructs Hurricane Milton at mainland Florida landfall near Siesta Key / Sarasota (Cat 3, 120 mph). It visualizes the storm's intense cyclonic rainbands, counter-clockwise rotation, and the resulting eyewall wind field that unleashed 100+ mph gusts and a 16.03-inch deluge directly across Tampa Bay.

```{=html}
<div class="news-chart-card" style="margin-bottom: 2rem;">
  <div style="display: flex; justify-content: space-between; align-items: center; margin-bottom: 0.75rem; flex-wrap: wrap; gap: 8px;">
    <div>
      <span class="live-pill" style="background: rgba(220, 38, 38, 0.2); color: #f87171; border: 1px solid #ef4444;"><span class="live-pulse-dot" style="background-color: #ef4444;"></span> REAL-TIME ROTATION SIMULATOR</span>
      <h3 style="color: #ffffff; margin: 0.35rem 0 0 0; font-size: 1.25rem; font-weight: 700;">Hurricane Milton: Animated Cloud Formation & Cyclonic Wind Field</h3>
      <p style="color: #94a3b8; font-size: 0.85rem; margin: 2px 0 0 0;">Interactive 60fps canvas simulation: counter-clockwise cyclonic spin of convective cloud bands and wind field points with inward surface inflow deflection.</p>
    </div>
    <div style="display: flex; gap: 8px; align-items: center;">
      <button id="milton-toggle-spin" class="live-refresh-btn" style="padding: 0.35rem 0.75rem; font-size: 0.82rem;">⏸ Pause</button>
      <button id="milton-speed-spin" class="live-refresh-btn" style="padding: 0.35rem 0.75rem; font-size: 0.82rem;">⚡ 2x Speed</button>
    </div>
  </div>
  <div style="position: relative; width: 100%; border-radius: 8px; overflow: hidden; background: #0b1320; border: 1px solid #334155;">
    <canvas id="milton-cyclone-canvas" class="cyclone-simulator-canvas" data-storm="milton" style="width: 100%; height: 560px; display: block;"></canvas>
  </div>
</div>
```

---

::: {.callout-note}
### Explore More Weather Crises
Return to the [Natural Disasters Overview](disasters.qmd) to view the full 10-year timeline, comparison tables, and summaries of all major storms.
:::
