---
title: "Interactive Weather Explorer"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
library(plotly)
library(leaflet)
library(jsonlite)
library(htmltools)

# Fetch initial real-time telemetry from NWS API for KTPA
live_init <- tryCatch({
  req <- fromJSON("https://api.weather.gov/stations/KTPA/observations/latest")
  p <- req$properties
  list(
    temp = paste0(round(p$temperature$value * 9/5 + 32, 1), "°F"),
    dew = paste0(round(p$dewpoint$value * 9/5 + 32, 1), "°F"),
    press = paste0(round(p$barometricPressure$value / 3386.39, 2), " inHg"),
    wind = if (round(p$windSpeed$value * 0.621371, 1) == 0) "Calm" else paste0(round(p$windSpeed$value * 0.621371, 1), " mph"),
    precip = if (is.null(p$precipitationLastHour$value) || is.na(p$precipitationLastHour$value)) "0.00 in" else paste0(round(p$precipitationLastHour$value * 39.3701, 2), " in"),
    condition = if (!is.null(p$textDescription)) p$textDescription else "Observed",
    time = format(as.POSIXct(p$timestamp), "%I:%M %p")
  )
}, error = \(e) {
  list(temp = "75.2°F", dew = "71.6°F", press = "29.93 inHg", wind = "Calm", precip = "0.00 in", condition = "Clear", time = "Live")
})

# Load and combine all 10 years of hourly data (2015–2024)
files <- list.files("data", pattern = "^ktpa_20[0-9]{2}\\.csv$", full.names = TRUE)

daily_weather <- map(files, \(f) {
  read_csv(f, col_types = cols(.default = col_character()), show_col_types = FALSE)
}) |>
  list_rbind() |>
  mutate(
    date = as.Date(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    dew_point = parse_number(HourlyDewPointTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  ) |>
  group_by(date) |>
  summarize(
    temp_avg = round(mean(temp, na.rm = TRUE), 1),
    temp_max = suppressWarnings(max(temp, na.rm = TRUE)),
    temp_min = suppressWarnings(min(temp, na.rm = TRUE)),
    dew_point_avg = round(mean(dew_point, na.rm = TRUE), 1),
    pressure_min = suppressWarnings(min(pressure, na.rm = TRUE)),
    pressure_avg = round(mean(pressure, na.rm = TRUE), 2),
    wind_max = suppressWarnings(max(wind, na.rm = TRUE)),
    precip_total = round(sum(precip, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Tracked major Florida natural disasters (2015–2024)
disasters <- tribble(
  ~date, ~event, ~storm_desc,
  as.Date("2016-09-02"), "Hurricane Hermine", "Tropical gale and coastal storm surge",
  as.Date("2017-09-10"), "Hurricane Irma", "Cat 4 storm; 46 mph sustained wind, 28.69 inHg pressure plunge",
  as.Date("2018-10-10"), "Hurricane Michael", "Cat 5 Gulf transit; reverse negative surge in Tampa Bay",
  as.Date("2022-09-28"), "Hurricane Ian", "Cat 4/5 system; drained Hillsborough Bay before winds shifted",
  as.Date("2022-12-24"), "Christmas Arctic Freeze", "Decade-record polar plunge; KTPA dropped to 29.0°F",
  as.Date("2023-08-30"), "Hurricane Idalia", "Record high-tide saltwater inundation along Bayshore Blvd",
  as.Date("2024-08-04"), "Hurricane Debby", "All-time 24-hr rainfall record: 16.14 inches",
  as.Date("2024-09-26"), "Hurricane Helene", "Historic 6-7 ft storm surge across Tampa Bay coastline",
  as.Date("2024-10-09"), "Hurricane Milton", "Direct strike; 16.03 in deluge and eyewall pressure collapse"
)

disaster_pts <- inner_join(disasters, daily_weather, by = "date")

# All-time 10-year record peaks for each measurement
rec_temp <- tribble(
  ~date, ~val, ~peak_title, ~peak_desc,
  as.Date("2023-08-11"), 98.0, "10-Year Record High: 98.0°F", "Hottest temperature recorded at KTPA across the 2015–2024 decade",
  as.Date("2022-12-24"), 29.0, "10-Year Record Low: 29.0°F", "Coldest temperature recorded during the Christmas 2022 Arctic Freeze"
)

rec_dew <- tribble(
  ~date, ~val, ~peak_title, ~peak_desc,
  as.Date("2024-08-07"), 82.0, "10-Year Record Dew Point: 82.0°F", "Peak atmospheric moisture indicating oppressive tropical humidity"
)

rec_press <- tribble(
  ~date, ~val, ~peak_title, ~peak_desc,
  as.Date("2017-09-10"), 28.69, "All-Time Lowest Pressure: 28.69 inHg", "Eye of Hurricane Irma passing over the Tampa Bay telemetry network"
)

rec_wind <- tribble(
  ~date, ~val, ~peak_title, ~peak_desc,
  as.Date("2017-09-10"), 46.0, "10-Year Peak Sustained Wind: 46.0 mph", "Hurricane Irma tropical gale (peak gusts reached 66 mph)"
)

rec_precip <- tribble(
  ~date, ~val, ~peak_title, ~peak_desc,
  as.Date("2024-08-04"), 16.14, "All-Time Wettest Day: 16.14 in", "Historic 24-hour rainfall record dumped by Hurricane Debby"
)

# Background vertical dashed lines for disaster dates
event_shapes <- map(disasters$date, \(d) {
  list(
    type = "line",
    x0 = d,
    x1 = d,
    y0 = 0,
    y1 = 1,
    yref = "paper",
    line = list(color = "rgba(220, 38, 38, 0.4)", width = 1.3, dash = "dash")
  )
})

# Callout annotations highlighting the highest rating of each measurement
ann_temp <- list(
  list(
    x = as.Date("2023-08-11"), y = 98.0,
    text = "<b>🏆 10-Yr Record High: 98°F</b><br>Aug 11, 2023",
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#b91c1c", ax = 40, ay = -40,
    bgcolor = "#fef2f2", bordercolor = "#dc2626", borderpad = 5,
    font = list(size = 12, color = "#991b1b")
  ),
  list(
    x = as.Date("2022-12-24"), y = 29.0,
    text = "<b>❄️ Record Low: 29°F</b><br>Dec 24, 2022 (Freeze)",
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#0284c7", ax = 40, ay = 35,
    bgcolor = "#f0f9ff", bordercolor = "#0284c7", borderpad = 5,
    font = list(size = 12, color = "#0369a1")
  )
)

ann_dew <- list(
  list(
    x = as.Date("2024-08-07"), y = 82.0,
    text = "<b>🏆 10-Yr Record Dew Point: 82°F</b><br>Aug 7, 2024 (Tropical Muggy)",
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#047857", ax = 45, ay = -40,
    bgcolor = "#ecfdf5", bordercolor = "#059669", borderpad = 5,
    font = list(size = 12, color = "#065f46")
  )
)

ann_press <- list(
  list(
    x = as.Date("2017-09-10"), y = 28.69,
    text = "<b>🌪️ All-Time Lowest Pressure: 28.69 inHg</b><br>Sep 10, 2017 (Hurricane Irma Eye)",
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#1d4ed8", ax = 50, ay = 40,
    bgcolor = "#eff6ff", bordercolor = "#2563eb", borderpad = 5,
    font = list(size = 12, color = "#1e40af")
  )
)

ann_wind <- list(
  list(
    x = as.Date("2017-09-10"), y = 46.0,
    text = "<b>💨 Peak Sustained Wind: 46 mph</b><br>Sep 10, 2017 (Irma - 66 mph gust)",
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#7e22ce", ax = 50, ay = -40,
    bgcolor = "#faf5ff", bordercolor = "#9333ea", borderpad = 5,
    font = list(size = 12, color = "#6b21a8")
  )
)

ann_precip <- list(
  list(
    x = as.Date("2024-08-04"), y = 16.14,
    text = "<b>🌧️ All-Time Wettest Day: 16.14 in</b><br>Aug 4, 2024 (Hurricane Debby)",
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#0284c7", ax = -50, ay = -40,
    bgcolor = "#f0f9ff", bordercolor = "#0284c7", borderpad = 5,
    font = list(size = 12, color = "#0369a1")
  )
)
```

## 1. 10-Year Weather Telemetry Explorer (2015–2024)

Explore all five core meteorological measurements across the full 10-year observational record (2015–2024) at Tampa International Airport (KTPA). Use the dropdown menu in the upper-right corner of the graph to transition between each measurement, and use the range selector or bottom slider to zoom into any time period:

- **Red Diamond Markers (♦)** and dashed vertical guidelines: Mark when major hurricanes and tropical natural disasters struck.
- **Gold Star Markers (⭐)**: Highlight all-time 10-year record highs and record lows.

```{r}
#| label: fig-10yr-telemetry-dropdown
plot_ly() |>
  # 1. Temperature
  add_lines(
    data = daily_weather, x = ~date, y = ~temp_avg,
    name = "Daily Avg Temp (°F)",
    line = list(color = "#D84315", width = 1.3),
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#c2410c;\">🌡️ Temperature</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Daily Average:</b> ", temp_avg, " °F<br>",
      "<b>Daily High:</b> ", temp_max, " °F | <b>Daily Low:</b> ", temp_min, " °F"
    )
  ) |>
  add_markers(
    data = disaster_pts, x = ~date, y = ~temp_avg,
    name = "Disaster Event",
    marker = list(color = "#dc2626", size = 8, symbol = "diamond"),
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#dc2626;\">🚨 ", event, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Temp Reading:</b> ", temp_avg, " °F (High: ", temp_max, "°F, Low: ", temp_min, "°F)<br>",
      "<i>", storm_desc, "</i>"
    )
  ) |>
  add_markers(
    data = rec_temp, x = ~date, y = ~val,
    name = "10-Year Record High/Low",
    marker = list(color = "#f59e0b", size = 12, symbol = "star", line = list(color = "#b45309", width = 1.5)),
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#b45309;\">⭐ ", peak_title, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<i>", peak_desc, "</i>"
    )
  ) |>
  # 2. Dew Point
  add_lines(
    data = daily_weather, x = ~date, y = ~dew_point_avg,
    name = "Avg Dew Point (°F)",
    line = list(color = "#00695C", width = 1.3),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#047857;\">💧 Dew Point Moisture</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Avg Dew Point:</b> ", dew_point_avg, " °F"
    )
  ) |>
  add_markers(
    data = disaster_pts, x = ~date, y = ~dew_point_avg,
    name = "Disaster Event",
    marker = list(color = "#dc2626", size = 8, symbol = "diamond"),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#dc2626;\">🚨 ", event, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Dew Point:</b> ", dew_point_avg, " °F<br>",
      "<i>", storm_desc, "</i>"
    )
  ) |>
  add_markers(
    data = rec_dew, x = ~date, y = ~val,
    name = "10-Year Record High/Low",
    marker = list(color = "#f59e0b", size = 12, symbol = "star", line = list(color = "#b45309", width = 1.5)),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#b45309;\">⭐ ", peak_title, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<i>", peak_desc, "</i>"
    )
  ) |>
  # 3. Barometric Pressure
  add_lines(
    data = daily_weather, x = ~date, y = ~pressure_min,
    name = "Lowest Station Pressure (inHg)",
    line = list(color = "#1565C0", width = 1.3),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#1d4ed8;\">🌪️ Barometric Pressure</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Minimum Pressure:</b> ", pressure_min, " inHg<br>",
      "<b>Daily Average:</b> ", pressure_avg, " inHg"
    )
  ) |>
  add_markers(
    data = disaster_pts, x = ~date, y = ~pressure_min,
    name = "Disaster Event",
    marker = list(color = "#dc2626", size = 8, symbol = "diamond"),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#dc2626;\">🚨 ", event, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Lowest Pressure:</b> ", pressure_min, " inHg<br>",
      "<i>", storm_desc, "</i>"
    )
  ) |>
  add_markers(
    data = rec_press, x = ~date, y = ~val,
    name = "10-Year Record High/Low",
    marker = list(color = "#f59e0b", size = 12, symbol = "star", line = list(color = "#b45309", width = 1.5)),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#b45309;\">⭐ ", peak_title, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<i>", peak_desc, "</i>"
    )
  ) |>
  # 4. Maximum Sustained Wind
  add_lines(
    data = daily_weather, x = ~date, y = ~wind_max,
    name = "Max Sustained Wind (mph)",
    line = list(color = "#6A1B9A", width = 1.3),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#7e22ce;\">💨 Wind Velocity</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Peak Sustained Wind:</b> ", wind_max, " mph"
    )
  ) |>
  add_markers(
    data = disaster_pts, x = ~date, y = ~wind_max,
    name = "Disaster Event",
    marker = list(color = "#dc2626", size = 8, symbol = "diamond"),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#dc2626;\">🚨 ", event, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Peak Sustained Wind:</b> ", wind_max, " mph<br>",
      "<i>", storm_desc, "</i>"
    )
  ) |>
  add_markers(
    data = rec_wind, x = ~date, y = ~val,
    name = "10-Year Record High/Low",
    marker = list(color = "#f59e0b", size = 12, symbol = "star", line = list(color = "#b45309", width = 1.5)),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#b45309;\">⭐ ", peak_title, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<i>", peak_desc, "</i>"
    )
  ) |>
  # 5. Precipitation & Deluges
  add_lines(
    data = daily_weather, x = ~date, y = ~precip_total,
    name = "Daily Rainfall Total (in)",
    line = list(color = "#0288D1", width = 1.3),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#0284c7;\">🌧️ Hydrological Telemetry</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Daily Rainfall:</b> ", precip_total, " inches"
    )
  ) |>
  add_markers(
    data = disaster_pts, x = ~date, y = ~precip_total,
    name = "Disaster Event",
    marker = list(color = "#dc2626", size = 8, symbol = "diamond"),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#dc2626;\">🚨 ", event, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Storm Rainfall:</b> ", precip_total, " inches<br>",
      "<i>", storm_desc, "</i>"
    )
  ) |>
  add_markers(
    data = rec_precip, x = ~date, y = ~val,
    name = "10-Year Record High/Low",
    marker = list(color = "#f59e0b", size = 12, symbol = "star", line = list(color = "#b45309", width = 1.5)),
    visible = FALSE,
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:14px; color:#b45309;\">⭐ ", peak_title, "</b><br>",
      "<b>📅 Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<i>", peak_desc, "</i>"
    )
  ) |>
  layout(
    title = list(
      text = "Tampa International Airport (KTPA) — Daily Temperatures (2015–2024)",
      font = list(size = 15, color = "#0f172a"),
      x = 0
    ),
    hovermode = "closest",
    hoverlabel = list(
      bgcolor = "rgba(255, 255, 255, 0.98)",
      bordercolor = "#475569",
      font = list(size = 14, family = "system-ui, -apple-system, sans-serif", color = "#0f172a"),
      align = "left"
    ),
    shapes = event_shapes,
    annotations = ann_temp,
    updatemenus = list(
      list(
        type = "dropdown",
        direction = "down",
        x = 1.0,
        y = 1.20,
        xanchor = "right",
        yanchor = "top",
        active = 0,
        bgcolor = "#ffffff",
        bordercolor = "#cbd5e1",
        font = list(size = 13, color = "#1e293b"),
        buttons = list(
          list(
            method = "update",
            label = "🌡️ Temperature (°F)",
            args = list(
              list(visible = list(TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)),
              list(
                title = list(text = "Tampa International Airport (KTPA) — Daily Temperatures (2015–2024)", font = list(size = 15, color = "#0f172a"), x = 0),
                yaxis = list(title = "Temperature (°F)", showgrid = TRUE),
                annotations = ann_temp
              )
            )
          ),
          list(
            method = "update",
            label = "💧 Dew Point Moisture (°F)",
            args = list(
              list(visible = list(FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)),
              list(
                title = list(text = "Tampa International Airport (KTPA) — Dew Point & Atmospheric Moisture (2015–2024)", font = list(size = 15, color = "#0f172a"), x = 0),
                yaxis = list(title = "Dew Point (°F)", showgrid = TRUE),
                annotations = ann_dew
              )
            )
          ),
          list(
            method = "update",
            label = "🌪️ Barometric Pressure (inHg)",
            args = list(
              list(visible = list(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)),
              list(
                title = list(text = "Tampa International Airport (KTPA) — Barometric Pressure (2015–2024)", font = list(size = 15, color = "#0f172a"), x = 0),
                yaxis = list(title = "Barometric Pressure (inHg)", showgrid = TRUE),
                annotations = ann_press
              )
            )
          ),
          list(
            method = "update",
            label = "💨 Max Sustained Wind (mph)",
            args = list(
              list(visible = list(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE, FALSE)),
              list(
                title = list(text = "Tampa International Airport (KTPA) — Maximum Sustained Wind Speeds (2015–2024)", font = list(size = 15, color = "#0f172a"), x = 0),
                yaxis = list(title = "Max Sustained Wind (mph)", showgrid = TRUE),
                annotations = ann_wind
              )
            )
          ),
          list(
            method = "update",
            label = "🌧️ Precipitation & Deluges (in)",
            args = list(
              list(visible = list(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE)),
              list(
                title = list(text = "Tampa International Airport (KTPA) — Daily Rainfall & Deluges (2015–2024)", font = list(size = 15, color = "#0f172a"), x = 0),
                yaxis = list(title = "Daily Rainfall (inches)", showgrid = TRUE),
                annotations = ann_precip
              )
            )
          )
        )
      )
    ),
    xaxis = list(
      title = "Date",
      showgrid = TRUE,
      rangeslider = list(visible = TRUE),
      rangeselector = list(
        x = 0, y = 1.20, xanchor = "left", yanchor = "top",
        buttons = list(
          list(count = 1, label = "1m", step = "month", stepmode = "backward"),
          list(count = 6, label = "6m", step = "month", stepmode = "backward"),
          list(count = 1, label = "1y", step = "year", stepmode = "backward"),
          list(count = 5, label = "5y", step = "year", stepmode = "backward"),
          list(step = "all", label = "All 10 Years")
        )
      )
    ),
    yaxis = list(title = "Temperature (°F)", showgrid = TRUE),
    margin = list(t = 70, b = 40, l = 55, r = 25)
  )
```

## 2. Tampa Bay Regional Weather Station Network

Click or hover on any station to view its elevation, telemetry type, and role in monitoring Tampa Bay's coastal weather extremes. The red marker highlights **KTPA**, the primary anchor station used in the 10-year analysis above.

```{r}
#| label: weather-station-map
get_station_obs <- function(station_id) {
  tryCatch({
    url <- paste0("https://api.weather.gov/stations/", station_id, "/observations/latest")
    req <- fromJSON(url)
    p <- req$properties
    tibble(
      id = station_id,
      live_temp = if (!is.null(p$temperature$value) && !is.na(p$temperature$value)) paste0(round(p$temperature$value * 9/5 + 32, 1), "°F") else "N/A",
      live_dew = if (!is.null(p$dewpoint$value) && !is.na(p$dewpoint$value)) paste0(round(p$dewpoint$value * 9/5 + 32, 1), "°F") else "N/A",
      live_press = if (!is.null(p$barometricPressure$value) && !is.na(p$barometricPressure$value)) paste0(round(p$barometricPressure$value / 3386.39, 2), " inHg") else "N/A",
      live_wind = if (!is.null(p$windSpeed$value) && !is.na(p$windSpeed$value)) {
        w <- round(p$windSpeed$value * 0.621371, 1)
        if (w == 0) "Calm" else paste0(w, " mph")
      } else "N/A",
      live_cond = if (!is.null(p$textDescription) && nzchar(p$textDescription)) p$textDescription else "Observed"
    )
  }, error = \(e) {
    tibble(id = station_id, live_temp = "N/A", live_dew = "N/A", live_press = "N/A", live_wind = "N/A", live_cond = "Offline")
  })
}

stations_raw <- tribble(
  ~id, ~name, ~county, ~lat, ~lon, ~elev_ft, ~type, ~description,
  "KTPA", "Tampa International Airport", "Hillsborough", 27.9755, -82.5332, 19, "Primary ASOS (Study Anchor)", "Central anchor station for this 10-year investigation. Recorded the 98.0°F record high and 16.14-inch Debby deluge.",
  "KPIE", "St. Pete–Clearwater International", "Pinellas", 27.9108, -82.6874, 11, "Primary ASOS", "Key coastal Pinellas station exposed directly to Old Tampa Bay marine breezes and tropical storm surge threats.",
  "KSPG", "Albert Whitted Airport", "Pinellas", 27.7651, -82.6270, 7, "ASOS", "Downtown St. Petersburg waterfront station; its low 7-foot elevation makes it critical for measuring bay water rise.",
  "KMCF", "MacDill Air Force Base", "Hillsborough", 27.8494, -82.5211, 14, "Military Aerodrome", "Located at the tip of the South Tampa peninsula between Hillsborough Bay and Old Tampa Bay.",
  "KVDF", "Tampa Executive Airport", "East Hillsborough", 28.0142, -82.3453, 22, "AWOS-3", "Inland Hillsborough station northeast of downtown Tampa; captures continental temperature swings away from water.",
  "KLAL", "Lakeland Linder International", "Polk", 27.9889, -82.0186, 142, "ASOS", "Inland station east of Tampa Bay and home base to the world-renowned NOAA Hurricane Hunters flight operations.",
  "KBKV", "Brooksville-Tampa Bay Regional", "Hernando", 28.4736, -82.4553, 77, "AWOS-3", "Northern perimeter station representing the Nature Coast boundary of the greater Tampa Bay region.",
  "KSRQ", "Sarasota Bradenton International", "Manatee / Sarasota", 27.3956, -82.5544, 30, "Primary ASOS", "Southern gateway station near the landfall impact zone of Hurricane Milton (2024)."
)

station_obs <- map(stations_raw$id, get_station_obs) |> list_rbind()

stations <- stations_raw |>
  left_join(station_obs, by = "id") |>
  mutate(
    marker_color = if_else(id == "KTPA", "#dc2626", "#2563eb"),
    fill_color   = if_else(id == "KTPA", "#ef4444", "#3b82f6"),
    radius_val   = if_else(id == "KTPA", 11, 8),
    live_snippet = paste0(
      "<div style=\"background:#f8fafc; border:1px solid #cbd5e1; border-radius:6px; padding:6px 8px; margin-top:8px;\">",
      "<span style=\"color:", if_else(id == "KTPA", "#dc2626", "#2563eb"), "; font-weight:700; font-size:11px;\">● LIVE STATION TELEMETRY</span><br>",
      "<b>Temp:</b> ", live_temp, " | <b>Dew:</b> ", live_dew, "<br>",
      "<b>Pressure:</b> ", live_press, " | <b>Wind:</b> ", live_wind, "<br>",
      "<b>Condition:</b> ", live_cond,
      "</div>"
    )
  )

station_popups <- paste0(
  "<div style=\"min-width:240px; font-family: system-ui, -apple-system, sans-serif;\">",
  "<h4 style=\"margin:0 0 6px 0; color:#1e3a8a; font-size:15px;\"><b>", stations$id, "</b> — ", stations$name, "</h4>",
  "<b>Network Type:</b> ", stations$type, "<br>",
  "<b>County:</b> ", stations$county, "<br>",
  "<b>Elevation:</b> ", stations$elev_ft, " ft<br>",
  "<b>Coordinates:</b> ", stations$lat, "° N, ", abs(stations$lon), "° W<br>",
  "<hr style=\"margin:8px 0; border:0; border-top:1px solid #cbd5e1;\">",
  "<p style=\"margin:0 0 4px 0; font-size:12.5px; color:#334155; line-height:1.4;\">", stations$description, "</p>",
  stations$live_snippet,
  "</div>"
)

leaflet(stations, height = 500) |>
  addProviderTiles(providers$Esri.WorldStreetMap) |>
  fitBounds(
    lng1 = min(stations$lon) - 0.14, lat1 = min(stations$lat) - 0.08,
    lng2 = max(stations$lon) + 0.14, lat2 = max(stations$lat) + 0.08
  ) |>
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = ~radius_val,
    color = ~marker_color,
    fillColor = ~fill_color,
    fillOpacity = 0.85,
    weight = 2,
    popup = station_popups,
    label = lapply(station_popups, htmltools::HTML),
    labelOptions = labelOptions(
      style = list("white-space" = "normal", "max-width" = "280px"),
      direction = "auto"
    )
  )
```
