---
title: "Bay Area Weather Analysis"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(leaflet)
library(jsonlite)
library(htmltools)

# Fetch initial real-time telemetry from NWS API for KTPA
live_init <- tryCatch({
  req <- fromJSON("https://api.weather.gov/stations/KTPA/observations/latest")
  p <- req$properties
  list(
    temp = paste0(round(p$temperature$value * 9/5 + 32, 1), "°F"),
    dew = paste0(round(p$dewpoint$value * 9/5 + 32, 1), "°F"),
    press = paste0(round(p$barometricPressure$value / 3386.39, 2), " inHg"),
    wind = if (round(p$windSpeed$value * 0.621371, 1) == 0) "Calm" else paste0(round(p$windSpeed$value * 0.621371, 1), " mph"),
    precip = if (is.null(p$precipitationLastHour$value) || is.na(p$precipitationLastHour$value)) "0.00 in" else paste0(round(p$precipitationLastHour$value * 39.3701, 2), " in"),
    condition = if (!is.null(p$textDescription)) p$textDescription else "Observed",
    time = format(as.POSIXct(p$timestamp), "%I:%M %p")
  )
}, error = \(e) {
  list(temp = "75.2°F", dew = "71.6°F", press = "29.93 inHg", wind = "Calm", precip = "0.00 in", condition = "Clear", time = "Live")
})
```

## Real-Time Telemetry Feed (KTPA)

Live observations streamed directly from the National Weather Service automated sensor network at Tampa International Airport.

```{=html}
<div class="live-telemetry-box">
  <div class="live-telemetry-header">
    <div style="display: flex; align-items: center; flex-wrap: wrap; gap: 10px;">
      <span class="live-pill"><span class="live-pulse-dot"></span> LIVE NWS BROADCAST</span>
      <span style="font-size: 1.15rem; font-weight: 700; color: #ffffff;">KTPA — Tampa International Airport Telemetry</span>
      <span id="live-time-str" style="color: #94a3b8; font-size: 0.85rem; font-weight: 500;">Observed Live</span>
    </div>
    <div style="display: flex; gap: 8px; align-items: center;">
      <span id="live-condition-badge" style="background: rgba(59, 130, 246, 0.2); border: 1px solid #3b82f6; color: #93c5fd; padding: 0.35rem 0.75rem; border-radius: 6px; font-size: 0.85rem; font-weight: 700;">Loading...</span>
      <button id="live-refresh-btn" class="live-refresh-btn" onclick="fetchLiveKTPAWeather()">🔄 Refresh</button>
    </div>
  </div>

  <div class="live-grid">
    <div class="live-card">
      <div class="live-card-icon">🌡️</div>
      <div id="live-temp-val" class="live-card-val">--.-°F</div>
      <div class="live-card-label">Live Temperature</div>
      <div class="live-card-sub">Measures ambient air warmth in the shade.</div>
    </div>
    <div class="live-card">
      <div class="live-card-icon">💧</div>
      <div id="live-dew-val" class="live-card-val">--.-°F</div>
      <div class="live-card-label">Live Dew Point</div>
      <div class="live-card-sub">Measures air moisture; higher values feel muggy.</div>
    </div>
    <div class="live-card">
      <div class="live-card-icon">🌪️</div>
      <div id="live-press-val" class="live-card-val">--.-- inHg</div>
      <div class="live-card-label">Live Pressure</div>
      <div class="live-card-sub">Barometric weight; falling pressure signals storms.</div>
    </div>
    <div class="live-card">
      <div class="live-card-icon">💨</div>
      <div id="live-wind-val" class="live-card-val">-- mph</div>
      <div class="live-card-label">Live Wind Speed</div>
      <div class="live-card-sub">Sustained horizontal airflow across airport runways.</div>
    </div>
    <div class="live-card">
      <div class="live-card-icon">🌧️</div>
      <div id="live-precip-val" class="live-card-val">-- in</div>
      <div class="live-card-label">Live Rain (1-Hr)</div>
      <div class="live-card-sub">Liquid rainfall accumulated over the past 60 minutes.</div>
    </div>
  </div>
</div>
```

## Tampa Bay Regional Weather Station Network

Click or hover on any station to view its elevation, telemetry type, and role in monitoring Tampa Bay's coastal weather extremes. The red marker highlights **KTPA**, the primary anchor station.

```{r}
#| label: weather-station-map
get_station_obs <- function(station_id) {
  tryCatch({
    url <- paste0("https://api.weather.gov/stations/", station_id, "/observations/latest")
    req <- fromJSON(url)
    p <- req$properties
    tibble(
      id = station_id,
      live_temp = if (!is.null(p$temperature$value) && !is.na(p$temperature$value)) paste0(round(p$temperature$value * 9/5 + 32, 1), "°F") else "N/A",
      live_dew = if (!is.null(p$dewpoint$value) && !is.na(p$dewpoint$value)) paste0(round(p$dewpoint$value * 9/5 + 32, 1), "°F") else "N/A",
      live_press = if (!is.null(p$barometricPressure$value) && !is.na(p$barometricPressure$value)) paste0(round(p$barometricPressure$value / 3386.39, 2), " inHg") else "N/A",
      live_wind = if (!is.null(p$windSpeed$value) && !is.na(p$windSpeed$value)) {
        w <- round(p$windSpeed$value * 0.621371, 1)
        if (w == 0) "Calm" else paste0(w, " mph")
      } else "N/A",
      live_cond = if (!is.null(p$textDescription) && nzchar(p$textDescription)) p$textDescription else "Observed"
    )
  }, error = \(e) {
    tibble(id = station_id, live_temp = "N/A", live_dew = "N/A", live_press = "N/A", live_wind = "N/A", live_cond = "Offline")
  })
}

stations_raw <- tribble(
  ~id, ~name, ~county, ~lat, ~lon, ~elev_ft, ~type, ~description,
  "KTPA", "Tampa International Airport", "Hillsborough", 27.9755, -82.5332, 19, "Primary ASOS (Study Anchor)", "Central anchor station for this 10-year investigation. Recorded the 98.0°F record high and 16.14-inch Debby deluge.",
  "KPIE", "St. Pete–Clearwater International", "Pinellas", 27.9108, -82.6874, 11, "Primary ASOS", "Key coastal Pinellas station exposed directly to Old Tampa Bay marine breezes and tropical storm surge threats.",
  "KSPG", "Albert Whitted Airport", "Pinellas", 27.7651, -82.6270, 7, "ASOS", "Downtown St. Petersburg waterfront station; its low 7-foot elevation makes it critical for measuring bay water rise.",
  "KMCF", "MacDill Air Force Base", "Hillsborough", 27.8494, -82.5211, 14, "Military Aerodrome", "Located at the tip of the South Tampa peninsula between Hillsborough Bay and Old Tampa Bay.",
  "KVDF", "Tampa Executive Airport", "East Hillsborough", 28.0142, -82.3453, 22, "AWOS-3", "Inland Hillsborough station northeast of downtown Tampa; captures continental temperature swings away from water.",
  "KLAL", "Lakeland Linder International", "Polk", 27.9889, -82.0186, 142, "ASOS", "Inland station east of Tampa Bay and home base to the world-renowned NOAA Hurricane Hunters flight operations.",
  "KBKV", "Brooksville-Tampa Bay Regional", "Hernando", 28.4736, -82.4553, 77, "AWOS-3", "Northern perimeter station representing the Nature Coast boundary of the greater Tampa Bay region.",
  "KSRQ", "Sarasota Bradenton International", "Manatee / Sarasota", 27.3956, -82.5544, 30, "Primary ASOS", "Southern gateway station near the landfall impact zone of Hurricane Milton (2024)."
)

station_obs <- map(stations_raw$id, get_station_obs) |> list_rbind()

stations <- stations_raw |>
  left_join(station_obs, by = "id") |>
  mutate(
    marker_color = if_else(id == "KTPA", "#dc2626", "#2563eb"),
    fill_color   = if_else(id == "KTPA", "#ef4444", "#3b82f6"),
    radius_val   = if_else(id == "KTPA", 11, 8),
    live_snippet = paste0(
      "<div style=\"background:#f8fafc; border:1px solid #cbd5e1; border-radius:6px; padding:6px 8px; margin-top:8px;\">",
      "<span style=\"color:", if_else(id == "KTPA", "#dc2626", "#2563eb"), "; font-weight:700; font-size:11px;\">● LIVE STATION TELEMETRY</span><br>",
      "<b>Temp:</b> ", live_temp, " | <b>Dew:</b> ", live_dew, "<br>",
      "<b>Pressure:</b> ", live_press, " | <b>Wind:</b> ", live_wind, "<br>",
      "<b>Condition:</b> ", live_cond,
      "</div>"
    )
  )

station_popups <- paste0(
  "<div style=\"min-width:240px; font-family: system-ui, -apple-system, sans-serif;\">",
  "<h4 style=\"margin:0 0 6px 0; color:#1e3a8a; font-size:15px;\"><b>", stations$id, "</b> — ", stations$name, "</h4>",
  "<b>Network Type:</b> ", stations$type, "<br>",
  "<b>County:</b> ", stations$county, "<br>",
  "<b>Elevation:</b> ", stations$elev_ft, " ft<br>",
  "<b>Coordinates:</b> ", stations$lat, "° N, ", abs(stations$lon), "° W<br>",
  "<hr style=\"margin:8px 0; border:0; border-top:1px solid #cbd5e1;\">",
  "<p style=\"margin:0 0 4px 0; font-size:12.5px; color:#334155; line-height:1.4;\">", stations$description, "</p>",
  stations$live_snippet,
  "</div>"
)

leaflet(stations, height = 500) |>
  addProviderTiles(providers$Esri.WorldStreetMap) |>
  fitBounds(
    lng1 = min(stations$lon) - 0.14, lat1 = min(stations$lat) - 0.08,
    lng2 = max(stations$lon) + 0.14, lat2 = max(stations$lat) + 0.08
  ) |>
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = ~radius_val,
    color = ~marker_color,
    fillColor = ~fill_color,
    fillOpacity = 0.85,
    weight = 2,
    popup = station_popups,
    label = lapply(station_popups, htmltools::HTML),
    labelOptions = labelOptions(
      style = list("white-space" = "normal", "max-width" = "280px"),
      direction = "auto"
    )
  )
```
