---
title: "Recent Data"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

::: {.news-header}
<span class="news-badge">CONTINUOUS TELEMETRY STREAM (2025 – PRESENT)</span>
<h1 style="color: #ffffff; margin-top: 0.5rem; margin-bottom: 0.5rem; font-size: 2.2rem; font-weight: 800;">Recent Weather Data</h1>
<div class="news-station-tag">STATION: TAMPA INTERNATIONAL AIRPORT (KTPA) &bull; 2025 OBSERVATIONS & LIVE STREAM</div>
:::

```{r}
#| label: setup-recent
library(tidyverse)
library(plotly)

recent_hourly <- read_csv("data/ktpa_2025_present.csv", col_types = cols(.default = col_character()), show_col_types = FALSE) |>
  mutate(
    time = as.POSIXct(DATE, format = "%Y-%m-%dT%H:%M:%S"),
    date = as.Date(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    dew = parse_number(HourlyDewPointTemperature),
    press = parse_number(HourlyStationPressure),
    wind = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  )

recent_daily <- recent_hourly |>
  group_by(date) |>
  summarize(
    wind = suppressWarnings(max(wind, na.rm = TRUE)),
    precip = round(sum(precip, na.rm = TRUE), 2),
    .groups = "drop"
  )
```

## Real-Time Telemetry Feed (KTPA)

Live surface meteorological observations streamed directly from the National Weather Service automated sensor network at Tampa International Airport.

```{=html}
<div class="live-telemetry-box">
  <div class="live-telemetry-header">
    <div style="display: flex; align-items: center; flex-wrap: wrap; gap: 10px;">
      <span class="live-pill"><span class="live-pulse-dot"></span> LIVE NWS BROADCAST</span>
      <span style="font-size: 1.15rem; font-weight: 700; color: #ffffff;">KTPA — Tampa International Airport Telemetry</span>
      <span id="live-time-str" style="color: #94a3b8; font-size: 0.85rem; font-weight: 500;">Observed Live</span>
    </div>
    <div style="display: flex; gap: 8px; align-items: center;">
      <span id="live-condition-badge" style="background: rgba(59, 130, 246, 0.2); border: 1px solid #3b82f6; color: #93c5fd; padding: 0.35rem 0.75rem; border-radius: 6px; font-size: 0.85rem; font-weight: 700;">Loading...</span>
      <button id="live-refresh-btn" class="live-refresh-btn" onclick="fetchLiveKTPAWeather()">🔄 Refresh</button>
    </div>
  </div>

  <div class="live-grid">
    <div class="live-card">
      <div class="live-card-icon">🌡️</div>
      <div id="live-temp-val" class="live-card-val">--.-°F</div>
      <div class="live-card-label">Live Temperature</div>
      <div class="live-card-sub">Measures ambient air warmth in the shade.</div>
    </div>
    <div class="live-card">
      <div class="live-card-icon">💧</div>
      <div id="live-dew-val" class="live-card-val">--.-°F</div>
      <div class="live-card-label">Live Dew Point</div>
      <div class="live-card-sub">Measures air moisture; higher values feel muggy.</div>
    </div>
    <div class="live-card">
      <div class="live-card-icon">🌪️</div>
      <div id="live-press-val" class="live-card-val">--.-- inHg</div>
      <div class="live-card-label">Live Pressure</div>
      <div class="live-card-sub">Barometric weight; falling pressure signals storms.</div>
    </div>
    <div class="live-card">
      <div class="live-card-icon">💨</div>
      <div id="live-wind-val" class="live-card-val">-- mph</div>
      <div class="live-card-label">Live Wind Speed</div>
      <div class="live-card-sub">Sustained horizontal airflow across airport runways.</div>
    </div>
    <div class="live-card">
      <div class="live-card-icon">🌧️</div>
      <div id="live-precip-val" class="live-card-val">-- in</div>
      <div class="live-card-label">Live Rain (1-Hr)</div>
      <div class="live-card-sub">Liquid rainfall accumulated over the past 60 minutes.</div>
    </div>
  </div>
</div>
```

---

## Continuous KTPA Telemetry & Trend Stream (2025 – Present)

This unified graph tracks all 5 meteorological measurements for KTPA from January 1, 2025 to the present. Use the **dropdown menu** below to switch between measurements, and the **range selector** (24h, 7d, 1m, 6m, 1y, All) or bottom slider to zoom into any time window.

```{r}
#| label: live-stream-2025-graph
plot_ly(elementId = "ktpa-2025-live-chart") |>
  # 1. Temperature (Trace 0)
  add_lines(
    data = recent_hourly, x = ~time, y = ~temp,
    name = "Hourly Temp (°F)",
    visible = TRUE,
    line = list(color = "#D84315", width = 1.3),
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:13px; color:#c2410c;\">🌡️ Temperature</b><br>",
      "<b>Time:</b> ", format(time, "%B %d, %Y %I:%M %p"), "<br>",
      "<b>Temp:</b> ", temp, " °F"
    )
  ) |>
  # 2. Dew Point (Trace 1)
  add_lines(
    data = recent_hourly, x = ~time, y = ~dew,
    name = "Hourly Dew Point (°F)",
    visible = FALSE,
    line = list(color = "#00695C", width = 1.3),
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:13px; color:#047857;\">💧 Dew Point</b><br>",
      "<b>Time:</b> ", format(time, "%B %d, %Y %I:%M %p"), "<br>",
      "<b>Dew Point:</b> ", dew, " °F"
    )
  ) |>
  # 3. Barometric Pressure (Trace 2)
  add_lines(
    data = recent_hourly, x = ~time, y = ~press,
    name = "Hourly Pressure (inHg)",
    visible = FALSE,
    line = list(color = "#1565C0", width = 1.3),
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:13px; color:#1d4ed8;\">🌪️ Barometric Pressure</b><br>",
      "<b>Time:</b> ", format(time, "%B %d, %Y %I:%M %p"), "<br>",
      "<b>Pressure:</b> ", press, " inHg"
    )
  ) |>
  # 4. Wind Speed (Trace 3)
  add_lines(
    data = recent_daily, x = ~date, y = ~wind,
    name = "Daily Max Wind (mph)",
    visible = FALSE,
    line = list(color = "#6A1B9A", width = 1.3),
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:13px; color:#7e22ce;\">💨 Max Wind Speed</b><br>",
      "<b>Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Max Wind:</b> ", wind, " mph"
    )
  ) |>
  # 5. Precipitation (Trace 4)
  add_lines(
    data = recent_daily, x = ~date, y = ~precip,
    name = "Daily Precip Total (in)",
    visible = FALSE,
    line = list(color = "#0288D1", width = 1.3),
    hoverinfo = "text",
    text = ~paste0(
      "<b style=\"font-size:13px; color:#0284c7;\">🌧️ Precipitation</b><br>",
      "<b>Date:</b> ", format(date, "%B %d, %Y"), "<br>",
      "<b>Daily Total:</b> ", precip, " in"
    )
  ) |>
  layout(
    title = list(
      text = "KTPA Continuous Meteorological Telemetry (2025 – Present & Live Stream)",
      font = list(size = 15, color = "#0f172a"),
      x = 0
    ),
    hovermode = "closest",
    hoverlabel = list(
      bgcolor = "rgba(255, 255, 255, 0.98)",
      bordercolor = "#475569",
      font = list(size = 13, family = "system-ui, -apple-system, sans-serif", color = "#0f172a"),
      align = "left"
    ),
    xaxis = list(
      title = "Timestamp",
      showgrid = TRUE,
      rangeslider = list(visible = TRUE),
      rangeselector = list(
        buttons = list(
          list(count = 1, label = "24h", step = "day", stepmode = "backward"),
          list(count = 7, label = "7d", step = "day", stepmode = "backward"),
          list(count = 1, label = "1m", step = "month", stepmode = "backward"),
          list(count = 6, label = "6m", step = "month", stepmode = "backward"),
          list(count = 1, label = "1y", step = "year", stepmode = "backward"),
          list(step = "all", label = "All (2025–Now)")
        )
      )
    ),
    yaxis = list(title = "Temperature (°F)", showgrid = TRUE),
    margin = list(t = 60, b = 40, l = 55, r = 25),
    updatemenus = list(
      list(
        type = "dropdown",
        direction = "down",
        x = 0.0,
        y = 1.18,
        xanchor = "left",
        yanchor = "bottom",
        buttons = list(
          list(
            method = "update",
            args = list(
              list(visible = c(TRUE, FALSE, FALSE, FALSE, FALSE)),
              list(yaxis = list(title = "Temperature (°F)", showgrid = TRUE))
            ),
            label = "🌡️ 1. Temperature (°F)"
          ),
          list(
            method = "update",
            args = list(
              list(visible = c(FALSE, TRUE, FALSE, FALSE, FALSE)),
              list(yaxis = list(title = "Dew Point (°F)", showgrid = TRUE))
            ),
            label = "💧 2. Dew Point (°F)"
          ),
          list(
            method = "update",
            args = list(
              list(visible = c(FALSE, FALSE, TRUE, FALSE, FALSE)),
              list(yaxis = list(title = "Barometric Pressure (inHg)", showgrid = TRUE))
            ),
            label = "🌪️ 3. Barometric Pressure (inHg)"
          ),
          list(
            method = "update",
            args = list(
              list(visible = c(FALSE, FALSE, FALSE, TRUE, FALSE)),
              list(yaxis = list(title = "Wind Speed (mph)", showgrid = TRUE))
            ),
            label = "💨 4. Sustained Wind Speed (mph)"
          ),
          list(
            method = "update",
            args = list(
              list(visible = c(FALSE, FALSE, FALSE, FALSE, TRUE)),
              list(yaxis = list(title = "Precipitation (inches)", showgrid = TRUE))
            ),
            label = "🌧️ 5. Precipitation (in)"
          )
        )
      )
    )
  )
```
