---
title: "Anatomy of a Storm"
format:
  html:
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-story
library(tidyverse)
library(patchwork)

# Load 10-year KTPA data
files <- list.files("data", pattern = "\\.csv$", full.names = TRUE)

weather <- map(files, \(f) {
  read_csv(f, col_types = cols(.default = col_character()), show_col_types = FALSE)
}) |>
  list_rbind() |>
  filter(grepl(":53:", DATE)) |>
  mutate(
    datetime = as.POSIXct(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  )

# 72-hour window centered on Hurricane Irma's closest approach
irma <- weather |>
  filter(datetime >= as.POSIXct("2017-09-09 12:00:00") & datetime <= as.POSIXct("2017-09-12 12:00:00"))
```

::: {.news-header}
<span class="news-badge">METEOROLOGICAL CASE STUDY</span>
<h1 style="color: #ffffff; margin-top: 0.5rem; margin-bottom: 0.5rem; font-size: 2.2rem; font-weight: 800;">Anatomy of a Hurricane Strike</h1>
<div class="news-station-tag">STATION: KTPA &bull; 72-HOUR HURRICANE IRMA TELEMETRY BREAKDOWN</div>
:::

## What Happens When a Hurricane Strikes Tampa?

In our 10-year meteorological record of Tampa International Airport (KTPA), one anomaly stands above all others: the early morning hours of September 11, 2017. As Hurricane Irma tore northward up the Florida peninsula, the airport's automated sensors recorded the violent physical signature of a tropical cyclone passing just miles to the east. By examining hourly barometric pressure, sustained wind speed, and precipitation together across this 72-hour window, we can observe the exact mechanics of what happens when a tropical cyclone strikes Tampa Bay.

```{r}
#| label: fig-irma-anatomy
#| fig-cap: "Hourly meteorological telemetry recorded at Tampa International Airport (KTPA) during the 72-hour passage of Hurricane Irma (September 9–12, 2017). Top: Station barometric pressure (inHg). Middle: Sustained wind speed (mph). Bottom: Hourly precipitation (inches)."
#| fig-width: 9.5
#| fig-height: 7.2
#| dpi: 180

# Panel 1: Barometric Pressure
p1 <- ggplot(irma, aes(x = datetime, y = pressure)) +
  geom_line(color = "#1d4ed8", linewidth = 1.1) +
  geom_point(data = filter(irma, pressure == min(pressure, na.rm = TRUE)), color = "#dc2626", size = 3) +
  annotate(
    "label", x = as.POSIXct("2017-09-11 02:00:00"), y = 28.78,
    label = "Lowest Pressure: 28.69 inHg\n(Eye Passage - 2:00 AM Sep 11)",
    fill = "#eff6ff", color = "#1e40af", fontface = "bold", size = 3.3
  ) +
  scale_y_continuous(breaks = seq(28.5, 30.5, by = 0.5), limits = c(28.5, 30.3)) +
  labs(
    title = "The Anatomy of a Hurricane Strike: Hurricane Irma at KTPA",
    subtitle = "Synchronized 72-Hour Telemetry at Tampa International Airport (Sep 9–12, 2017)",
    y = "Pressure (inHg)",
    x = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 14, color = "#0f172a"),
    plot.subtitle = element_text(color = "#475569", size = 11, margin = margin(b = 10)),
    axis.title.y = element_text(face = "bold", color = "#1e40af", size = 10),
    panel.grid.minor = element_blank()
  )

# Panel 2: Wind Speed
p2 <- ggplot(irma, aes(x = datetime, y = wind)) +
  geom_line(color = "#7e22ce", linewidth = 1.1) +
  geom_point(data = filter(irma, wind == max(wind, na.rm = TRUE)), color = "#dc2626", size = 3) +
  annotate(
    "label", x = as.POSIXct("2017-09-11 02:00:00"), y = 37,
    label = "Peak Sustained Wind: 41+ mph\n(Squall Gusts to 66 mph)",
    fill = "#faf5ff", color = "#6b21a8", fontface = "bold", size = 3.3
  ) +
  scale_y_continuous(breaks = seq(0, 50, by = 10), limits = c(0, 48)) +
  labs(y = "Wind Speed (mph)", x = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    axis.title.y = element_text(face = "bold", color = "#6b21a8", size = 10),
    panel.grid.minor = element_blank()
  )

# Panel 3: Hourly Rainfall
p3 <- ggplot(irma, aes(x = datetime, y = precip)) +
  geom_col(fill = "#0284c7", width = 3200) +
  scale_y_continuous(breaks = seq(0, 1.5, by = 0.5), limits = c(0, 1.2)) +
  labs(
    y = "Rain (in/hr)",
    x = "Date & Local Time (EDT)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.title.x = element_text(face = "bold", color = "#0f172a", margin = margin(t = 8)),
    axis.title.y = element_text(face = "bold", color = "#0369a1", size = 10),
    panel.grid.minor = element_blank()
  )

# Combine using patchwork
p1 / p2 / p3
```

### The Barometric Engine

Under ordinary conditions, atmospheric pressure across Tampa Bay hovers around a stable **30.10 inHg** (1,019 hPa). As Irma’s circulation approached the Florida Keys on September 9, KTPA’s barometer began a steady, steep descent. By 2:00 AM on September 11, the barometer plummeted to a decade-low record of **28.69 inHg** (971.5 hPa). This precipitous V-shaped drop created a steep pressure gradient across western Florida, essentially turning the skies above Tampa into a giant atmospheric siphon that drew in gale-force winds from hundreds of miles away.

### Wind and Rain in Lockstep

The middle and bottom panels reveal how closely winds and rainfall sync with barometric collapse. As the pressure bottomed out during the early hours of September 11, sustained winds surged from calm 5–10 mph sea breezes into tropical storm-force gales of **41 to 46 mph**, with violent squall gusts recorded up to **66 mph**. Simultaneously, heavy spiral rain bands dumped torrential precipitation over Hillsborough County, dropping over **4.5 inches of rain** in just a few concentrated hours before the storm center moved north toward Georgia.

### The Broader Lesson: Why Storm Track Dictates Tampa Bay's Fate

What makes Hurricane Irma’s signature particularly revealing is how track geography dictates Tampa Bay’s vulnerability. Because Irma tracked just inland to the east of Tampa, the storm’s counterclockwise rotation drove surface winds out of the northeast and east. These offshore winds pushed bay water outward into the Gulf of Mexico, causing a dramatic **"negative storm surge"** that temporarily drained the bay floor instead of flooding coastal neighborhoods. 

By contrast, when **Hurricane Helene** (September 2024) passed offshore in the Gulf to the west, and **Hurricane Milton** (October 2024) made landfall just south near Siesta Key, their counterclockwise winds blew onshore into Tampa Bay. That onshore push, combined with historic rainfall (Milton dumped 16.03 inches at KTPA), transformed that same atmospheric pressure collapse into devastating coastal storm surge. The data confirms that while barometric drop dictates storm intensity, it is the storm's track relative to Tampa Bay that determines whether water is drained away or driven ashore.

---

### Comparative Cyclone Telemetry at Tampa (KTPA)

| Storm / Metric | Minimum Pressure | Peak Sustained Wind | Peak Gust | Total Storm Deluge | Coastal Surge Outcome |
|:---|:---|:---|:---|:---|:---|
| **Hurricane Irma** (Sep 2017) | **28.69 inHg** (10-Yr Record Low) | 46.0 mph | 66.0 mph | 4.57 in | Negative Surge (Bay drained) |
| **Hurricane Helene** (Sep 2024) | 29.56 inHg | 38.0 mph | 55.0 mph | 2.14 in | Record Coastal Surge (6–7 ft flood) |
| **Hurricane Milton** (Oct 2024) | 29.18 inHg | 30.0+ mph | 60.0+ mph | **16.03 in** | Severe Inland Flooding & Eyewall Rain |

---

## Hurricane Irma Landfall: Shape, Rotation & Wind Circulation

The visualization below illustrates Hurricane Irma at the precise moment of its mainland Florida landfall at **Marco Island** (September 10, 2017 at 3:35 PM EDT as a Category 3 hurricane with 115 mph winds). It reveals the storm's characteristic cyclonic shape, counter-clockwise rotation, and the resulting surface wind directions that siphoned water away from Tampa Bay.

```{=html}
<div class="news-chart-card" style="margin-bottom: 2rem;">
  <div style="display: flex; justify-content: space-between; align-items: center; margin-bottom: 0.75rem; flex-wrap: wrap; gap: 8px;">
    <div>
      <span class="live-pill" style="background: rgba(220, 38, 38, 0.2); color: #f87171; border: 1px solid #ef4444;"><span class="live-pulse-dot" style="background-color: #ef4444;"></span> REAL-TIME ROTATION SIMULATOR</span>
      <h3 style="color: #ffffff; margin: 0.35rem 0 0 0; font-size: 1.25rem; font-weight: 700;">Hurricane Irma: Animated Cloud Formation & Cyclonic Wind Field</h3>
      <p style="color: #94a3b8; font-size: 0.85rem; margin: 2px 0 0 0;">Interactive 60fps canvas simulation: counter-clockwise cyclonic spin of convective cloud bands and wind field points with inward surface inflow deflection.</p>
    </div>
    <div style="display: flex; gap: 8px; align-items: center;">
      <button id="irma-toggle-spin" class="live-refresh-btn" style="padding: 0.35rem 0.75rem; font-size: 0.82rem;">⏸ Pause</button>
      <button id="irma-speed-spin" class="live-refresh-btn" style="padding: 0.35rem 0.75rem; font-size: 0.82rem;">⚡ 2x Speed</button>
    </div>
  </div>
  <div style="position: relative; width: 100%; border-radius: 8px; overflow: hidden; background: #0b1320; border: 1px solid #334155;">
    <canvas id="irma-cyclone-canvas" style="width: 100%; height: 560px; display: block;"></canvas>
  </div>
</div>
```

```{r}
#| label: fig-irma-landfall-map
#| fig-cap: "Meteorological reconstruction of Hurricane Irma at mainland landfall (Marco Island, FL). Vectors illustrate the counter-clockwise surface wind field (with a ~20° inward inflow angle), highlighting how offshore northeasterly gales over Tampa Bay generated an historic negative surge while driving severe onshore flooding into Southwest Florida."
#| fig-width: 9.5
#| fig-height: 7.5
#| dpi: 180

# Load Florida coastline and county outlines
fl_states <- ggplot2::map_data("state") |> filter(region %in% c("florida", "georgia", "alabama"))
fl_counties <- ggplot2::map_data("county") |> filter(region == "florida")

# Fixed Landfall Center: Marco Island, FL
landfall_lon <- -81.72
landfall_lat <- 25.94

# 1st Florida Landfall: Cudjoe Key
cudjoe_lon <- -81.38
cudjoe_lat <- 24.70

# Station Location: Tampa International Airport (KTPA)
tampa_lon <- -82.53
tampa_lat <- 27.98

# Historical Track Points
track_df <- tibble(
  lon = c(-81.38, -81.72, -81.82, -81.88, -82.02, -82.18, -82.35),
  lat = c(24.70, 25.94, 26.50, 27.10, 27.75, 28.35, 29.20)
)

# Concentric Wind Radii (Eyewall, Hurricane-Force 74+ mph, Tropical Storm 39+ mph)
radii_df <- tibble(
  radius_deg = c(0.38, 0.95, 2.35),
  zone = factor(c("Eyewall (115 mph)", "Hurricane-Force (74+ mph)", "Tropical Storm Force (39+ mph)"),
                levels = c("Eyewall (115 mph)", "Hurricane-Force (74+ mph)", "Tropical Storm Force (39+ mph)"))
) |>
  reframe(
    tibble(
      angle = seq(0, 2 * pi, length.out = 150),
      lon = landfall_lon + radius_deg * cos(angle) * 1.15,
      lat = landfall_lat + radius_deg * sin(angle)
    ),
    .by = c(zone, radius_deg)
  )

# Spiral Rainband Arms (Logarithmic spiral geometry)
spiral_bands <- purrr::map(c(0, pi/2, pi, 3*pi/2), \(offset) {
  theta <- seq(0.35, 4.3, length.out = 100)
  r <- 0.25 * exp(0.56 * theta)
  tibble(
    lon = landfall_lon + r * cos(-theta + offset) * 1.15,
    lat = landfall_lat + r * sin(-theta + offset),
    arm = as.character(offset)
  )
}) |> list_rbind()

# Irma Convective Cloud Puffs along spiral arms
cloud_puffs <- purrr::map(c(0, pi/2, pi, 3*pi/2), \(offset) {
  theta <- seq(0.35, 4.2, length.out = 50)
  r <- 0.25 * exp(0.56 * theta)
  tibble(
    lon = landfall_lon + r * cos(-theta + offset) * 1.15,
    lat = landfall_lat + r * sin(-theta + offset),
    size = seq(4, 13, length.out = 50)
  )
}) |> list_rbind()

# Cyclonic Wind Vector Grid across Florida
grid_pts <- expand_grid(
  lon = seq(-85.4, -79.2, by = 0.50),
  lat = seq(24.2, 30.4, by = 0.50)
) |>
  mutate(
    dx = (lon - landfall_lon) / 1.15,
    dy = lat - landfall_lat,
    dist = sqrt(dx^2 + dy^2)
  ) |>
  filter(dist >= 0.35 & dist <= 3.9) |>
  mutate(
    phi = atan2(dy, dx),
    wind_dir = phi + pi/2 + 0.36, # Counter-clockwise tangential + 20 deg surface friction inflow
    arrow_len = case_when(
      dist < 0.95 ~ 0.32,
      dist < 2.35 ~ 0.27,
      TRUE ~ 0.21
    ),
    u = arrow_len * cos(wind_dir) * 1.15,
    v = arrow_len * sin(wind_dir),
    wind_intensity = case_when(
      dist < 0.95 ~ "Hurricane Force (74+ mph)",
      dist < 2.35 ~ "Tropical Storm Force (39-73 mph)",
      TRUE ~ "Outer Gale (< 39 mph)"
    )
  )

ggplot() +
  # Regional land geography
  geom_polygon(data = fl_states, aes(x = long, y = lat, group = group),
               fill = "#1e293b", color = "#475569", linewidth = 0.45) +
  geom_polygon(data = fl_counties, aes(x = long, y = lat, group = group),
               fill = NA, color = "#334155", linewidth = 0.25) +
  # Irma Central Dense Overcast (CDO) Cloud Shield
  geom_point(aes(x = landfall_lon, y = landfall_lat), size = 30, color = "#f8fafc", alpha = 0.20) +
  geom_point(aes(x = landfall_lon, y = landfall_lat), size = 18, color = "#ffffff", alpha = 0.28) +
  # Spiral Convective Cloud Puffs
  geom_point(data = cloud_puffs, aes(x = lon, y = lat, size = size),
             color = "#f8fafc", alpha = 0.16) +
  scale_size_identity() +
  # Tropical Storm & Hurricane Wind Radii
  geom_polygon(data = filter(radii_df, zone == "Tropical Storm Force (39+ mph)"),
               aes(x = lon, y = lat), fill = "#f59e0b", alpha = 0.12, color = "#f59e0b", linetype = "dotted") +
  geom_polygon(data = filter(radii_df, zone == "Hurricane-Force (74+ mph)"),
               aes(x = lon, y = lat), fill = "#ef4444", alpha = 0.18, color = "#ef4444", linetype = "dashed") +
  geom_polygon(data = filter(radii_df, zone == "Eyewall (115 mph)"),
               aes(x = lon, y = lat), fill = "#991b1b", alpha = 0.30, color = "#dc2626") +
  # Spiral Rainband Arms
  geom_path(data = spiral_bands, aes(x = lon, y = lat, group = arm),
            color = "#38bdf8", alpha = 0.40, linewidth = 1.3) +
  # Wind Direction Vectors (Color-Coded by Intensity)
  geom_segment(data = grid_pts,
               aes(x = lon, y = lat, xend = lon + u, yend = lat + v, color = wind_intensity),
               arrow = arrow(length = unit(0.18, "cm"), type = "closed"),
               linewidth = 0.55, alpha = 0.85) +
  scale_color_manual(
    name = "Wind Field Velocity",
    values = c(
      "Hurricane Force (74+ mph)" = "#ef4444",
      "Tropical Storm Force (39-73 mph)" = "#38bdf8",
      "Outer Gale (< 39 mph)" = "#64748b"
    )
  ) +
  # Historical Irma Track Line
  geom_path(data = track_df, aes(x = lon, y = lat),
            color = "#f8fafc", linewidth = 1.2, linetype = "longdash") +
  geom_point(data = track_df, aes(x = lon, y = lat),
             color = "#f8fafc", size = 2) +
  # Fixed Landfall Marker (Marco Island)
  geom_point(aes(x = landfall_lon, y = landfall_lat),
             shape = 21, size = 6, fill = "#dc2626", color = "#ffffff", stroke = 2) +
  annotate(
    "label", x = landfall_lon + 0.35, y = landfall_lat - 0.20,
    label = "MAINLAND LANDFALL\nMarco Island (Cat 3, 115 mph)\nSep 10, 2017 • 3:35 PM EDT",
    hjust = 0, fill = "#7f1d1d", color = "#ffffff", fontface = "bold", size = 3.2,
    label.padding = unit(0.3, "lines"), label.r = unit(0.2, "lines")
  ) +
  # Cudjoe Key Initial Landfall Marker
  geom_point(aes(x = cudjoe_lon, y = cudjoe_lat),
             shape = 21, size = 4, fill = "#e11d48", color = "#ffffff", stroke = 1.2) +
  annotate(
    "label", x = cudjoe_lon - 0.20, y = cudjoe_lat - 0.32,
    label = "1st FL Landfall: Cudjoe Key (Cat 4, 130 mph)",
    hjust = 0.5, fill = "#1e1b4b", color = "#f43f5e", fontface = "bold", size = 2.8,
    label.padding = unit(0.2, "lines")
  ) +
  # Tampa Bay / KTPA Marker & Offshore Wind Callout
  geom_point(aes(x = tampa_lon, y = tampa_lat),
             shape = 21, size = 5.5, fill = "#0284c7", color = "#ffffff", stroke = 1.8) +
  annotate(
    "label", x = -83.1, y = 28.5,
    label = "TAMPA BAY (KTPA)\nOffshore NE Gale Winds\nNegative Surge Drained Bay Bed\nPressure: 28.69 inHg • Gusts: 66 mph",
    hjust = 0.5, fill = "#082f49", color = "#38bdf8", fontface = "bold", size = 3.2,
    label.padding = unit(0.3, "lines"), label.r = unit(0.2, "lines")
  ) +
  # Counter-Clockwise Rotation Callout
  annotate(
    "label", x = -84.3, y = 26.2,
    label = "COUNTER-CLOCKWISE ROTATION\n• Surface Inflow Angle ~20°\n• Siphoned Water Out of Tampa Bay\n• Drove Onshore Surge in SW Florida",
    fill = "#030d17", color = "#e2e8f0", fontface = "bold", size = 3.1,
    label.padding = unit(0.35, "lines"), label.r = unit(0.2, "lines")
  ) +
  coord_quickmap(xlim = c(-85.8, -79.2), ylim = c(23.8, 30.4)) +
  labs(
    title = "Hurricane Irma Landfall Dynamics & Cyclonic Wind Circulation",
    subtitle = "Fixed Mainland Landfall (Marco Island, FL) • Cyclonic Spiral Rainbands & Surface Wind Field",
    caption = "NOAA / NWS / KTPA station records. Vectors denote surface wind direction showing how offshore NE gales siphoned water out of Tampa Bay.",
    x = "Longitude (°W)",
    y = "Latitude (°N)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.margin = margin(12, 20, 12, 20),
    plot.background = element_rect(fill = "#0b1320", color = NA),
    panel.background = element_rect(fill = "#0f172a", color = "#334155"),
    panel.grid.major = element_line(color = "#334155", linetype = "dotted"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(color = "#ffffff", face = "bold", size = 13, margin = margin(b = 4)),
    plot.subtitle = element_text(color = "#94a3b8", size = 8.8, margin = margin(b = 8)),
    plot.caption = element_text(color = "#64748b", size = 7.8, margin = margin(t = 8)),
    legend.position = "inside",
    legend.position.inside = c(0.82, 0.78),
    legend.background = element_rect(fill = "#0f172a", color = "#334155"),
    legend.title = element_text(color = "#ffffff", size = 8.5, face = "bold"),
    legend.text = element_text(color = "#cbd5e1", size = 7.8),
    axis.title = element_text(color = "#94a3b8", size = 9),
    axis.text = element_text(color = "#64748b", size = 8)
  )
```

### Key Takeaways from the Circulation Pattern:

1. **Fixed Landfall Anchor:** Irma struck Florida twice—first at **Cudjoe Key** as a Category 4 storm (130 mph), and second as a Category 3 mainland landfall on **Marco Island** (115 mph).
2. **Spiral Shape and Rainbands:** The logarithmic spiral bands depict the sprawling radius of the storm's outer rain shield, which extended across both the Gulf of Mexico and the Atlantic Ocean.
3. **Counter-Clockwise Rotation:** Like all Northern Hemisphere cyclones, winds rotate counter-clockwise around the low-pressure center, with surface friction deflecting winds slightly inward (~20° cross-isobar inflow).
4. **The Negative Surge Mechanism:** Because Tampa Bay was situated to the northwest of Irma's path, local surface winds blew from the **northeast (offshore)**. Instead of forcing Gulf waters onshore, these violent winds blew water out into the Gulf, exposing the seabed along Bayshore Boulevard.

